\documentclass[12pt]{article}
\usepackage{longtable}
\textwidth=7in
\textheight=9.5in
\topmargin=-1in
\headheight=0in
\headsep=.5in
\hoffset  -.85in

\pagestyle{empty}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\begin{document}

\begin{center}
{\bf CS165 Database Systems  
\\ DCS UP Diliman
\\ 2nd Semester, SY 2013-2014
\\ Sat 1:00 - 6:00 PM,  Room:  TBD  
}
\end{center}

\setlength{\unitlength}{1in}

\begin{picture}(6,.1) 
\put(0,0) {\line(1,0){6.25}}         
\end{picture}

 

\renewcommand{\arraystretch}{2}

\vskip.25in
\noindent\textbf{Instructor:} Michael Chua (cs165.13.14.2@gmail.com)
\vskip.25in
\noindent\textbf{Textbook:}  Database Management Systems, 3rd Edition (Ramakrishnan and Gehrke)

\vskip.25in
\noindent\textbf{Prerequisites:} CS32

\vskip.25in
\noindent \textbf{Course Objectives}:
\begin{itemize}
  \item Introduce fundamental concepts in data management and storage through database systems.
  \item Provide opportunities to apply these concepts through designing, implementing and manipulating database schemas.
  \item Learn to use an actual database management system.
  \item Explore modern alternatives to traditional database management systems.
\end{itemize}

\vspace*{.25in}
\noindent \textbf{Course Schedule:} 

\begin{center}
  \begin{longtable}{ |l|l|l|l| }
    \hline
    Week & Date & Lecture & Lab \\ \hline
    01 & Nov-16-2013 & 
    	Introduction and Database Design & 
    	Postgres Setup and Basics \\ \hline
    02 & Nov-23-2013 & 
    	The Relational Model &
    	SQL (Definitions and Manipulation) \\ \hline
    03 & Dec-07-2013 & 
    	Relational Algebra & 
    	SQL (Queries) \\ \hline
    04 & Dec-14-2013 & 
    	Relational Calculus &
    	SQL (Joins and Aggregations) \\ \hline
    05 & Jan-11-2013 & 
    	\multicolumn{2}{|c|}{First Exam} \\ \hline
    06 & Jan-18-2013 & 
    	Normalization & 
    	SQL (Functions and Triggers) \\ \hline
    07 & Jan-25-2013 & 
		Transaction Management &
		Django, Python and Postgres  \\ \hline
    08 & Feb-01-2013 & 
		Concurrency Control &
		Postgres Concurrency Control  \\ \hline
    09 & Feb-08-2013 & 
		NoSQL &
    	Django, Python and CouchDB \\ \hline
    10 & Feb-15-2013 & 
	    Storage and Indexing &
		Postgres Indexes \\ \hline
    11 & Feb-22-2013 &
	    Tree Indexing &
		? \\ \hline
    12 & Mar-01-2013 &
	    Hash Indexing &
		? \\ \hline
    13 & Mar-08-2013 &
		Security and Authorization &
    	Postgres Accounts \\ \hline
    14 & Mar-15-2013 & \multicolumn{2}{|c|}{Second Exam}\\
    \hline
  \end{longtable}
\end{center}

\noindent\textbf{Evaluation:} 
\begin{itemize}
  \item[] 25\% First Exam
  \item[] 25\% Second Exam
  \item[] 10\% Homework
  \item[] 20\% Report (NoSQL)
  \item[] 20\% Project
\end{itemize}

\vspace*{.25in}
\noindent\textbf{Grading}:

\begin{center}
  \begin{tabular}{ |l|l|l|l| }
    \hline
    General Average & Final Grade & General Average & Final Grade \\ \hline
    92-100 & 1.00 & 74-\textless78 & 2.25 \\ \hline
    88-\textless92 & 1.25 & 70-\textless74 & 2.50 \\ \hline
    85-\textless88 & 1.50 & 65-\textless70 & 2.75 \\ \hline
    82-\textless85 & 1.75 & 60-\textless65 & 3.00 \\ \hline
    78-\textless82 & 2.00 & \textless60 & 5.00 \\ \hline
  \end{tabular}
\end{center}

\iffalse
\vspace*{.25in}
\noindent\textbf{Class Policies:} 
\begin{itemize}
  \item Cheating will be punished according to university standards.
  \item Missed requirements are given a grade of 0. Excuses will be considered on a case-to-case basis.
  \item Use black or blue pens for written requirements.
  \item Attendance is not tracked. You can come in and leave at any time, but you're on your own for any activities or lectures you miss. 
  \item Avoid disruptive behavior.
  \item Put phones on silent. Vibrate is okay. Go outside room if you want to make or take a call.
  \item Contact me via e-mail if you wish to schedule a consultation. Please do it early.
  \item Feedback and criticism is welcomed.
\end{itemize}

\vspace*{.25in}

\vskip.25in
\noindent\textbf{Important Dates}:
\begin{center} \begin{minipage}{5in}
\begin{flushleft}
Bonifacio Day (Holiday) \dotfill Nov-30-2013\\
Christmas Break (Start) \dotfill Dec-19-2013\\
Christmas Break (End) \dotfill Jan-05-2013\\
First Exam \dotfill Jan-15-2014\\
Drop Deadline \dotfill Feb-20-2014\\
Second Exam  \dotfill Mar-15-2014\\
Report Deadline \dotfill Mar-30-2014\\
Project Deadline \dotfill Mar-30-2014\\
\end{flushleft}
\end{minipage}
\end{center}
\fi

\end{document}