\documentclass[12pt]{article}
\usepackage{longtable}
\textwidth=7in
\textheight=9.5in
\topmargin=-1in
\headheight=0in
\headsep=.5in
\hoffset  -.85in

\pagestyle{empty}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\begin{document}

\begin{center}
{\bf CS165 Database Systems  
\\ Homework 4 - Relational Calculus
}
\end{center}

\setlength{\unitlength}{1in}

\begin{picture}(6,.1) 
\put(0,0) {\line(1,0){6.25}}         
\end{picture}

\vskip.25in
\noindent\textbf{Submission:} Saturday December 21, 23:59:99, cs165.13.14.2@gmail.com

\vskip.25in
\noindent \textbf{Requirements}:
\begin{itemize}
  \item Send answers in pdf format. Strongly recommend use of LaTex. 
\end{itemize}

\vskip.25in
\noindent \textbf{Problem}:
\vskip.15in
\noindent For the same schema and queries given in homework 3, provide a correct TRC query for each odd numbered item. Also provide a correct DRC query for each even numbered item.

\begin{itemize}
	\item[] Suppliers(\underline{sid}: integer, sname: string, address: string)
	\item[] Parts(\underline{pid}: integer, pname: string, color: string)
 	\item[] Catalog(\underline{sid}: integer, \underline{pid}: integer, cost: real)
\end{itemize}

\begin{enumerate}
	\item Find the names of suppliers who supply some red part.
	\item Find the sids of suppliers who supply some red or green part.
	\item Find the sids of suppliers who supply some red part or are at 221 Packer Street. 
	\item Find the sids of suppliers who supply some red part and some green part. 
	\item Find the sids of suppliers who supply every part. 
	\item Find the sids of suppliers who supply every red part. 
	\item Find the sids of suppliers who supply every red or green part. 
	\item Find the sids of suppliers who supply every red part or supply every green part. 
	\item Find pairs of sids such that the supplier with the first sid charges more for some part than the supplier with the second sid. 
	\item Find the pids of parts supplied by at least two different suppliers. 
	\item Find the pids of the most expensive parts supplied by suppliers named Yosemite Sham. 
	\item Find the pids of parts supplied by every supplier at less than \$200. (If any supplier either does not supply the part or charges more than \$200 for it, the part is not selected.)
\end{enumerate}

\end{document}