\documentclass[10pt]{article}
\textwidth=7in
\textheight=9.5in
\topmargin=-1in
\headheight=0in
\headsep=.5in
\hoffset=-.85in
\pagestyle{empty}
\usepackage{amsmath}

\begin{document}

\begin{center}
{\bf CS165 Database Systems \\
Concurrency Control (Summary) }
\end{center}

\vskip.35in
\noindent {\bf Why?}

\vskip.10in
\noindent Multiple transactions running at the same time can cause records in the database to become inconsistent. The {\it scheduler} is responsible for arranging read and write actions to preserve consistency. It generates a {\it schedule}, which is an ordered list of actions for each transaction.

\vskip.35in
\noindent {\bf Serial and Serializable Schedules}

\vskip.10in
\noindent A schedule is {\it serial} if one transaction finishes completely before running the next transaction ($T_1$ first before $T_2$ or vice versa):

\begin{quote}
\begin{tabular}{l|l|l|l}
$T_1$			&$T_2$			&$A$	  &$B$	\\
\hline
 		    	&				& 25 	  & 25 	\\
$READ(A,t)$ 	& 				& 	 	  &    	\\
$t:=t+100$		&				&		  &	   	\\
$WRITE(A,t)$	&				& 125	  &	   	\\
$READ(B,t)$ 	& 				& 	 	  &    	\\
$t:=t+100$		&				&		  &	   	\\
$WRITE(B,t)$	&				& 		  & 125 \\
			 	&$READ(A,s)$ 	& 	 	  &    	\\
				&$s:=s*2$		&		  &	   	\\
				&$WRITE(A,s)$	& 250	  &		\\
				&$READ(B,s)$	&		  &		\\
				&$s:=s*2$		&		  &		\\
				&$WRITE(B,s)$	&		  & 250 \\
\end{tabular}
\end{quote}

\vskip.10in
\noindent Actions for multiple transactions can be {\it interleaved} (i.e., interrupt $T_1$ to let some actions of $T_2$ run). A schedule is {\it serializable} if the interleaved actions preserve consistency:

\begin{quote}
\begin{tabular}{l|l|l|l}
$T_1$			&$T_2$			&$A$	  &$B$	\\
\hline
 		    	&				& 25 	  & 25 	\\
$READ(A,t)$ 	& 				& 	 	  &    	\\
$t:=t+100$		&				&		  &	   	\\
$WRITE(A,t)$	&				& 125	  &	   	\\
				& $READ(A,s)$	&		  &		\\
				& $s:=s*2$		&		  &		\\
				& $WRITE(A,s)$  & 250	  &		\\
$READ(B,t)$		&				&		  &		\\
$t:=t+100$		&				&		  &		\\
$WRITE(B,t)$	&				&		  & 125 \\
				 $READ(B,s)$	&		  & 	\\
				& $s:=s*2$		&		  &		\\
				& $WRITE(B,s)$	&		  & 250 \\
\end{tabular}
\end{quote}

\vskip.10in
\noindent Not all schedules lend themselves to serializability. A schedule is {\it non-serializable} if the interleaved actions would break consistency:

\begin{quote}
\begin{tabular}{l|l|l|l}
$T_1$			&$T_2$			&$A$	  &$B$	\\
\hline
 		    	&				& 25 	  & 25 	\\
$READ(A,t)$ 	& 				& 	 	  &    	\\
$t:=t+100$		&				&		  &	   	\\
$WRITE(A,t)$	&				& 125	  &	   	\\
				& $READ(A,s)$	&		  &		\\
				& $s:=s*2$		&		  &		\\
				& $WRITE(A,s)$  & 250	  &		\\
				& $READ(B,s)$	&		  & 	\\
				& $s:=s*2$		&		  &		\\
				& $WRITE(B,s)$	&		  & 50  \\
$READ(B,t)$		&				&		  &		\\
$t:=t+100$		&				&		  &		\\
$WRITE(B,t)$	&				&		  & 150 \\
\end{tabular}
\end{quote}

\vskip.35in
\noindent {\bf A Notation for Transactions and Schedules}

\vskip.10in
\noindent We will concern ourselves with operations performed on the database objects, and not its values.

\vskip.10in
\noindent A transaction can be represented as sequence of read and write actions:

\begin{quote}
$T_1: r_1(A); w_1(A); r_1(B); w_1(B);$
\vskip.10in
$T_2: r_2(A); w_2(A); r_2(B); w_2(B);$
\end{quote}

\vskip.10in
\noindent An {\it action} is an expression of the form $r_i(X)$ or $w_i(X)$. \\

\noindent A {\it schedule} $S$ is a long chain of actions, which may have interleaving transactions:

\begin{quote}
$S: r_1(A); w_1(A); r_1(B); w_1(B);r_2(A); w_2(A); r_2(B); w_2(B);$
\end{quote}

\vskip.15in
\noindent {\bf Conflict-Serializability}

\vskip.10in
\noindent Our goal is to test a schedule's serializability by swapping some of its adjacent actions. 

\vskip.10in
\noindent A {\it conflict} is when two pairs of actions, when swapped, led to a change in a transaction's result.

\vskip.10in
\noindent Our goal would be to test a schedule's serializability by swapping some of its adjacent actions belonging to different transactions. If a schedule claims to be conflict-serializable, then it can be transformed into a serial schedule.

\vskip.10in
\noindent A conflict occurs if (1) we swap two actions of the same transaction, or (2) swap two actions that access the same object (with at least one write action).

\vskip.15in
\noindent {\bf Precedence Graphs}

\vskip.10in
\noindent When conflicting pairs of actions appear in a schedule, its sequence of transactions must appear in the same order. $T_1$ {\it takes precedence over} $T_2$ if there are actions $A_1$ of $T_1$ and $A_2$ of $T_2$, such that:

\vskip.10in
\begin{quote}
1. $A_1$ is ahead of $A_2$ in $S_1$. \\
2. Both $A_1$ and $A_2$ involve the same object. \\
3. At least one of $A_1$ and $A_2$ is a write action.
\end{quote}

\vskip.10in
\noindent Construct a graph where each node is a transaction. An edge exists from $T_1$ to $T_2$ if some action in $T_1$ conflicts with a later action in $T_2$. A schedule is conflict-serializable if and only if the precedence graph is acyclic.

\vskip.15in
\noindent {\bf Locking}

\vskip.10in
\noindent A step towards ensuring serializability is to add a lock to an object. While locked, other transactions do not have access to the object. Once the transaction is finished with the object, it unlocks it.

\vskip.15in
\noindent {\bf Two-Phase Locking}

\vskip.10in
\noindent All transactions in a schedule first request locks (acquire phase), and then release these locks (release phase).

\vskip.15in
\noindent {\bf Deadlock}

\vskip.10in
\noindent While two-phase locking prevents transactions for requesting locks after a specific time, it still does not solve the problem when transactions are made to wait for others to release their locks.

\vskip.15in
\noindent {\bf Shared and Exclusive Locks}

\vskip.10in
\noindent One problem with locks is that a transaction is forced to acquire a lock on an object even if it only needs to read the object. This increases the chances of deadlock. One trade-off is to have two locks with different strengths (write lock stronger than read lock).

\vskip.10in
\noindent Shared ($sl_i(X)$) and Exclusive ($xl_i(X)$) locks have three requirements:

\vskip.10in
\noindent 1. {\it Consistency of transactions}. A transaction may not write without holding an exclusive lock, and you may not read without holding some lock. More precisely, in any transaction $T_i$,

\begin{quote}
(a) A read action $r_i(X)$ must be preceded by $sl_i(X)$ or $xl_i(X)$, with no intervening $u_i(X)$. \\
(b) A write action $w_i(X)$ must be preceded by $xl_i(X)$, with no intervening $u_i(X)$.
\end{quote}

\vskip.10in
\noindent All locks must be followed by an unlock of the same object.

\vskip.10in
\noindent 2. {\it Two-phase locking of transactions}. Locking must precede unlocking. To be more precise, in any two-phased locked transaction $T_i$, no action $sl_i(X)$ or $xl_i(X)$ can be preceded by an action $u_i(Y)$, for any $Y$.

\vskip.10in
\noindent 3. {\it Legality of schedules}. An object may either be locked exclusively by one transaction or by several in shared mode, but not both. More precisely:

\begin{quote}
(a) If $xl_i(X)$ appears in a schedule, then there cannot be a following $xl_j(X)$ or $sl_j(X)$, for some $j$ other than $i$, without an intervening $ul_i(X)$. \\
(b) If $sl_i(X)$ appears in a schedule, then there cannot be a following $xl_j(X)$, for some $j$ other than $i$, without an intervening $ul_i(X)$.
\end{quote}

\vskip.10in
\noindent It is possible for a transaction to have both shared and exclusive locks on the same object, provided it does not conflict with the lock(s) of other transactions.

\vskip.35in
\noindent {\bf Reference}
\vskip.10in
Database Systems - The Complete Book (Garcia-Molina, Ullman, Widom)

\end{document}