\documentclass[12pt]{article}
\usepackage{longtable}
\usepackage{hyperref}
\textwidth=7in
\textheight=9.5in
\topmargin=-1in
\headheight=0in
\headsep=.5in
\hoffset  -.85in

\pagestyle{empty}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\begin{document}

\begin{center}
{\bf CS165 Database Systems  
\\ Written Report - Survey on NoSQL Technologies
}
\end{center}

\setlength{\unitlength}{1in}

\begin{picture}(6,.1) 
\put(0,0) {\line(1,0){6.25}}         
\end{picture}

\vskip.25in
\noindent \textbf{Overview}: \\
This paper is intended as an opportunity for students to explore current modern database system technologies that fall beyond the scope of the traditional material presented in the textbook and lectures. It also serves as practice for critical writing, which will be essential in your last year's dissertation, and for both graduate studies and most professions. You will work on this paper in the same groups that you are in for the project. If any of you wish to work in any other grouping, please contact us through the class e-mail as soon as possible. 

\vskip.25in
\noindent \textbf{Requirements}: \\
The topic of this paper will be \textbf{NoSQL}. NoSQL refers to a loosely-related collection of database technologies that have gained prominence over the past five years. These technologies are all attempts to provide alternatives to the traditional relational model that might be more responsive to the present-day needs of modern applications. 

NoSQL is still a very young topic by the standards of our field. A commonly-used taxonomy classifies NoSQL into the following groups:

\begin{itemize}
\item Key-Value Stores
\item Document Stores
\item Column Stores
\end{itemize}

Your task will be to write an \textbf{introductory survey} that covers \textbf{one} of these topics, which will be \textbf{assigned} to you. The overall goal of your paper should be to broadly introduce your topic to someone familiar with traditional databases, but not to NoSQL. 

Your paper should be at least \textbf{8 pages} long, and no more than \textbf{10 pages}, excluding the bibliography. Your writing should have supporting references, and should attribute to them properly. \textbf{Avoid plagiarism}. 

\vskip.25in
\noindent \textbf{Evaluation}: \\
Your content and structure will be assessed on how well they answer the following questions:

\begin{itemize}
\item What are \textbf{essential characteristics} of your assigned NoSQL category, that sets it apart from traditional databases, and other NoSQL technologies? What makes it different?

\item What is the \textbf{history} of your assigned NoSQL category? What persons, organizations, companies and research have contributed to its development? Why is your category the way it is?

\item What are important \textbf{examples} of existing implementations of your category? What are they being used for? How are they being applied?

\item What are the \textbf{advantages} that your category provides that others do not? When should it be used?

\item What are the \textbf{drawbacks and limitations} of your category? When should it not be used?

\end{itemize}  
  
\vskip.25in
\noindent\textbf{Submission:} \\
A pdf copy of your paper should be submitted by e-mail on \textbf{March 30, 2014}. It is recommended that you use \LaTeX in writing your paper. An SCM like git or mercurial might also help you manage multiple versions of your document as you write it, and merging your collaborations together safely.

\end{document}